---
title: "Loken_HW5"
author: "Luke Loken"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
#Load options
  knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.path='Figs/',echo = TRUE)
```

```{r, include=FALSE}
#Load data and libraries
library(tidyr)
library(dplyr)
library(sp)
library(rgdal, verbose=F)
library(SDMTools)
library(raster,verbose = T)
library(RColorBrewer)
br <- brick('../../Data/Rasters/Complete_TAVG_LatLong1.nc')
br_WI <- brick('../../Data/Rasters/WI_CAVG_LatLong1.nc')
states <-  readOGR("../../Data/Shapefiles",  'StateOutlines')
states<-spTransform(states,  crs(br_WI))
WI<-states[states$NAME=='Wisconsin',]
ME<-readOGR("../../Data/Shapefiles", "Mendota_shape", verbose=F)
ME<-spTransform(ME, crs(br_WI))
```

**Dear Dr. Dugan:**

Use the 'Data/WI_CAVG_LatLong1.nc' file. In this lecture we dealt with the "temperature" variable. This file is the "climatology" variable.

Find the metadata for these dataset. http://berkeleyearth.org/data/. We are using Gridded Data. Monthly Land. Average Temperature (TAVG; 1753 – Recent).

## Question 1: The climatology variable is a monthly average for each cell. What years does this average represent?

Direct quote from (http://berkeleyearth.org), "climatology: For each grid cell, an estimate of the true surface temperature for each month during the period January 1951 to December 1980 reported in degrees C."

**Thus each climatology value is the average monthly temperature between 1951 and 1980**


## Question 2: Plot the August averages for Wisconsin.


```{r, fig1, fig.width = 5, echo=F}
# br_WI
August = br_WI[[8]] 
par(mgp=c(2,.5,0), tck=-.02)
plot(August, axes=T, col=brewer.pal(n = 12, name = "RdYlBu") , xlab=expression(paste('Longitude (', degree, ')')), ylab=expression(paste('Latitude (', degree, ')')))
plot(WI, add=T)
mtext('Mean August temperature',3,.5)
mtext(expression(paste('', degree, 'C')),4, 4.5)

```

## Extract the averages for the cell over Lake Mendota. Plot the monthly temperature averages.


```{r, include=T}
METemp = extract(br_WI,y = ME)
output = data.frame(Month = 1:12, Temp = METemp[[1]][1,])
row.names(output) <- month.abb[1:12]
```

```{r, fig2, fig.width = 6, echo=F}
par(mar=c(3,3,3,1),mgp=c(2,0.5,0), tck=-.01)
plot(output$Month,output$Temp,pch=16, type='o', xaxt='n', las=1, col=adjustcolor('red4',0.9), xlab='', ylab=expression(paste('Temperature (', degree, 'C)')))
abline(h=0, col='lightgrey', lty=3)
mtext('Monthly mean air temperatures (1951-1980)', 3, 1.5)
mtext('for Lake Mendota, WI', 3, 0.25)
axis(1, at=1:12, labels=month.abb[1:12])

```

## What is the August average for Lake Mendota?

```{r, include=T}
output[which(output$Month==8),]
```

**The average August air temperature over Lake Mendota is 20.19 degrees celcius.**
