---
title: "Loken_HW2"
author: "Luke Loken"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
  knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.path='Figs/',echo = TRUE)
```
  
```{r, include=FALSE}
library(sp)
library(rgdal)
library(raster)
ME<-readOGR(paste0("E:/Git_Repo/ZOO955/Data/Shapefiles"), "Mendota_shape")
NLC<-raster("E:/Git_Repo/ZOO955/Data/Rasters/nlcd_2011_landcover_2011_edition_2014_10_10/nlcd_2011_landcover_2011_edition_2014_10_10.img", package='raster')
```
  
Dear Dr. Dugan:

## Do you know what lake this is? 

```{r, fig1, fig.height = 5, fig.width = 5, echo=F}
source('MendotaMap.R')
```

Hint- it might be world famous. 


## What is the CRS for the 2011 National Land Cover dataset?


```{r, fig2, fig.height = 3.5, fig.width = 5, echo=F}
par(mar=c(3,3,0.5,0.5))
plot(NLC, axes=T)
# points(0,0,pch=10, cex=2, col='darkred')
# text(500000,0, 'Origin', col='darkred')
```
```{r, include=T}
NLCCRS<-crs(NLC)
print(NLCCRS)
```

Answer: USA Contiguous albers equal area, North American Datum 1983

## Determining if two spatial objects have the same CRS 

Use the `identical()` function


```{r}
identical(crs(ME), crs(CFL))
identical(crs(ME), crs(NLC))
```

Lake Mendota (ME) and the Center for Limnology (CFL) spatial objects have the same CRS.

Lake Mendota does not have the same CRS as the National Land Cover spatial object.




