---
title: "Loken_HW3"
author: "Hilary Dugan + Luke Loken"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
  knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.path='Figs/',echo = TRUE)
```
  
  # Hilary's answers
  
  ## Question 1
Define functions from the `sf` package. Good resource: https://en.wikipedia.org/wiki/DE-9IM

`st_intersects` =

`st_disjoint` =

`st_touches` =

`st_crosses` =

`st_within` =

`st_contains` =

`st_overlaps` =

`st_equals` =

`st_covers` =

`st_covered_by` =

`st_equals_exact` =

`st_is_within_distance` =

`st_buffer` =

`st_boundary` =

`st_convexhull` =

`st_union_cascaded` =

`st_simplify` =

`st_triangulate` =

`st_polygonize` =

`st_centroid` =

`st_segmentize` =

`st_union` =

## Question 2 
**Make a 500 m buffer of the 4 southern LTER lakes. Which buffers overlap?**

*Lakes Data*:
https://lter.limnology.wisc.edu/dataset/north-temperate-lakes-lter-yahara-lakes-district-boundary


```{r, include=T, warning=F, message=F}
library(sp)
library(rgdal)
library(raster)
library(sf)
library(tidyr)
library(dplyr)
Lakes_sp <- readOGR(paste0("E:/Git_Repo/ZOO955/Data/Shapefiles"), "yld_study_lakes", stringsAsFactors = F)
Lakes_sf <- st_read("E:/Git_Repo/ZOO955/Data/Shapefiles/yld_study_lakes.shp", stringsAsFactors = F)

#Make individual lake objects
ME <- Lakes_sf %>% filter(LAKEID == 'ME')
MO <- Lakes_sf %>% filter(LAKEID == 'MO')
WI <- Lakes_sf %>% filter(LAKEID == 'WI')
FI <- Lakes_sf %>% filter(LAKEID == 'FI')

#Transform to lat/long for axes plotting
Lakes_sf_latlong <- st_transform(Lakes_sf, crs=4326)

#Merging Hilary's and Luke's objects
lakes <- Lakes_sf

# What are the lake IDs?
ids <- lakes$LAKEID
names <- lakes$LAKE_NAME
```
  


Make 500 m buffer 
```{r, message=FALSE,warning=FALSE}
buffer500 = st_buffer(lakes,500)
#Check if buffers overlap
st_overlaps(buffer500)
```

We know the order of the lake IDS (FI, ME, MO, WI). So based on the overlap matrix. 

* Mendota overlaps with Monona. 
* Monona overlaps with Mendota and Wingra
* Wingra overlaps with Monona

**Or we can add lake ids to table**

```{r, message=FALSE,warning=FALSE}
# Repeat with sparse=False
overlaps <- st_overlaps(buffer500, sparse = FALSE)
colnames(overlaps) <- ids
rownames(overlaps) <- ids
print(overlaps)
```


## Question 3
**Increase the size of the lakes by 2x. What is the percent of Mendota that overlaps with Monona?**

Note: 

* You can't use a buffer because that does not retain the shape of the lakes. 
* You can't just multiply the lakes x2, because that multiplies the coordinates. You end up with the lakes somewhere other than Wisconsin. 

Instead:

* Find the distance from the edge of the lake to the centroid. Multiply these distances by 2. 

```{r, message=FALSE,warning=FALSE}
# Take just the geometry of the lakes 
glakes = st_geometry(lakes)
# Find the centroid
cntrd = st_centroid(glakes)

# Find distance from edge of lakes to centroid
cDist = (glakes - cntrd) 
# Multiply this distance by 2 and add back to centroid 
glakes2 = cDist * 2 + cntrd

# Find the intersection between Mendota and Monona. We know these are lakes 2 and 3. 
int = st_intersection(glakes2[2],glakes2[3])
# What is the size difference between the intersection and Mendota
st_area(int)/ st_area(glakes2[2])
```
**The percent of Mendota that overlaps with Monona is 19.4%**

If you want to double check by plotting 
```{r, message=FALSE,warning=FALSE}
# Plot glakes and centroid
plot(glakes,col='cadetblue')
plot(cntrd,col='red3',pch=16,add=T)
# Plot lakes double the size
plot(glakes2,border='red4',add=T)
# Plot intersection 
plot(int,col='red4',add=T)
```
  
  
  
  

Dear Dr. Dugan:

## Question 1: What does `st_buffer` do? 

```{r, fig1, fig.height = 5, fig.width = 6, echo=F}
ME_centroid<-st_centroid(ME)
ME_centroidbuff<-st_buffer(ME_centroid, dist=300)
ME_centroidbuffminus<-st_buffer(ME_centroid, dist=(-100))
ME_500buff<-st_buffer(ME, dist=500)
ME_500buffminus<-st_buffer(ME, dist=(-500))
plot(ME_500buff$geometry, lty=2, lwd=2)
plot(ME$geometry, add=T, col='lightblue', lwd=5)
plot(ME_500buffminus$geometry, add=T, col=NA,lty=2, lwd=2, border='red')
legend('topleft', c('Outer 500 m', 'Inner 500 m', 'Centroid 300 m'), bty='n', lty=c(2,2, 1), cex=1, lwd=2, title='Buffer', col=c('black', 'red', 'limegreen'))
plot(ME_centroidbuff$geometry, col=NA, add=T, lty=1, lwd=2, border='limegreen')
plot(ME_centroid$geometry, pch=18, col='mediumpurple4', add=T)
```

`st_buffer` creates a new polygon that all points are X distance away from the edge of another polygon, line or point. The buffer can extend outward (+) or inward (-) of a polygon.


## Question 2: Make a 500 m buffer around the 4 southern LTER lakes, which buffers overlap?

```{r, fig2, fig.height = 5.5, fig.width = 5.5, echo=F, warning=F}
Lakes_sf500<-st_buffer(Lakes_sf, 500)
MO_buff<-Lakes_sf500 %>% filter(LAKEID == 'MO')
Other_buff<-Lakes_sf500 %>% filter(LAKEID != 'MO')
intersection<-st_intersection(MO_buff, Other_buff)
plot(Lakes_sf500$geometry, graticule = st_crs(Lakes_sf_latlong), axes=T)
plot(Lakes_sf$geometry, add=T, col='lightblue')
plot(intersection$geometry, add=T, col='red')
box(which='plot')
```

```{r, fig3, fig.height = 6, fig.width = 6, echo=F, warning=F}
plot(intersection$geometry, col='red', graticule = st_crs(Lakes_sf_latlong), axes=T)
plot(MO_buff$geometry, axes = TRUE, add=T)
plot(Lakes_sf$geometry, add=T, col='lightblue')
plot(Lakes_sf500$geometry, add=T)
# plot(intersection$geometry, add=T, col='red')
box(which='plot')

```

Based on `st_overlaps`, the buffer around Lake Monona (ID#3) overlaps both the Lake Mendota (ID#2) and Lake Wingra (ID#4) buffers. No other polygons overlap.  

```{r}
overlaps<-st_overlaps(Lakes_sf500)
names(overlaps)<-Lakes_sf500$LAKE_NAME
str(overlaps)
```


## Question 3: Increase the size of the lakes by 2x, What percent of Mendota overlaps with Monona?

```{r, include=T, warning=F}
ME_centroid<-st_centroid(ME)
ME_points<-st_cast(ME, "POINT")

x<-st_coordinates(ME_points)[,1]
y<-st_coordinates(ME_points)[,2]

x2<-2*x-st_coordinates(ME_centroid)[,1]
y2<-2*y-st_coordinates(ME_centroid)[,2]

ME_points2<-st_multipoint(as.matrix(data.frame(x2, y2)))
ME_points3<-st_sfc(ME_points2)

ME_Lines<-st_multilinestring(ME_points3)
ME_Polygon<-st_polygonize(ME_Lines)

ME_Polygon2<-st_geometry(ME_Polygon, type = 3)
st_crs(ME_Polygon2) <- st_crs(ME)
overlap<-st_intersection(ME_Polygon2, MO)
```
```{r, fig4, fig.height = 5, fig.width = 5, echo=F, warning=F}
plot(ME_Polygon, col=NA, lwd=3)
plot(Lakes_sf$geometry, col='lightblue', add=T)
plot(ME_Polygon, col=NA, lwd=3, add=T)
plot(overlap, add=T, col='red')
```
```{r}
overlap_area<-st_area(overlap)
BigMendota_area<-st_area(ME_Polygon2)
as.numeric(overlap_area/BigMendota_area)
```

7.45% of the expanded Lake Mendota polygon overlaps Lake Monona. 



# More code to work through question 3

```{r, include=T, warning=F}
# Take just the geometry of the lakes 
glakes = st_geometry(lakes)
# Find the centroid
cntrd = st_centroid(glakes)

plot(glakes, col='lightblue')
plot(cntrd, add=T, pch=16)

# Find distance from edge of lakes to centroid
cDist = (glakes - cntrd) 
plot(cDist[c(2,3,4,1)], col=c('red', 'blue', 'yellow', 'plum3'), axes=T)
```

Note that `cntrd` is multiple polygon, where all polygons are centered at zero (centroid) and the x/y coordinates are distances **i.e., not coordinates** 

```{r, include=T, warning=F}
# Multiply this distance by 2 and add back to centroid 
glakes2 = cDist * 2 + cntrd

# Find the intersection between Mendota and Monona. We know these are lakes 2 and 3. 
int = st_intersection(glakes2[2],glakes2[3])
# What is the size difference between the intersection and Mendota
st_area(int)/ st_area(glakes2[2])
```

## Another way to do it, without creating a 'distance' polygon

```{r, include=T, warning=F}
# using geometry you can simply double the coordinates and subtract by the centroid
glakes3 = glakes*2 - cntrd

int = st_intersection(glakes3[2],glakes3[3])

st_area(int)/ st_area(glakes3[2])

plot(glakes3[2:3], col=c('blue', 'lightblue'))
plot(int, col=c('red'), add=T)
```

This does not work if you use the sf object. This code creates an error
`lakes3 <- lakes*2 - st_centroid(lakes)`

But if you use the `st_geometry` function you can do geometry math
```{r, include=T, warning=F}
biglakes <- st_geometry(lakes)*2 - st_centroid(st_geometry(lakes))
plot(biglakes, col=c('red', 'blue', 'yellow', 'plum3'))
plot(lakes, add=T, col=NA)
```