---
title: "Loken_HW4"
author: "Luke Loken"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
  knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.path='Figs/',echo = TRUE)
```

```{r, include=FALSE}
library(tidyr)
library(dplyr)
library(sp)
library(rgdal, verbose=F)
library(SDMTools)
library(raster,verbose = T)
nlcd = raster('../../Data/Rasters/MendotaNLCD.grd')
ME<-readOGR("../../Data/Shapefiles", "Mendota_shape", verbose=F)
ME2<-spTransform(ME, crs(nlcd))
```

Dear Dr. Dugan:

## Question 1: What is the percent of forest in the Mendota catchment?

```{r, fig1, fig.width = 4, echo=F}

nlcd_for <- nlcd
nlcd_for@legend@colortable[2:length(nlcd_for@legend@colortable)] = 'moccasin'
nlcd_for@legend@colortable[1] = 'white'
nlcd_for@legend@colortable[41:43] = 'darkgreen'

par(mar=rep(.25, 4), mfrow=c(1,1))
bbox<-bbox(nlcd)
plot(ME2, col='lightblue', border=NA, xlim=bbox[1,1:2], ylim=bbox[2,1:2], xaxs='i',  yaxs='i')
plot(nlcd_for, add=T)
plot(ME2, add=T, col='lightblue', border=NA)
legend("topleft", legend = c("Forest", "Non-forest watershed", "Lake Mendota"), fill = c('darkgreen', 'moccasin', 'lightblue'), bty='n')
box(which='plot')

usr<-par('usr')
Scalebar(x=usr[2]-15000, y=usr[3]+2000, distance=10000, scale=0.001, unit='km')  

```

```{r, include=T}

#Calculate with Lake Mendota included in catchment
nlcdVals = freq(nlcd)

# Sum values, minus zero values 
total = as.data.frame(nlcdVals) %>% dplyr::filter(value >= 1) %>%
  summarise_at('count',sum)

# Total forest counts
forest = as.data.frame(nlcdVals) %>% dplyr::filter(value >= 41 & value <= 43) %>%
  summarise_at('count',sum)

perFor = forest/total
perFor

#Remove Lake Mendota from catchment
#Mask raster by Lake Mendota
nlcd_nolake <- mask(nlcd, ME2, inverse=T)
nlcdVals_nolake = freq(nlcd_nolake)

total_nolake = as.data.frame(nlcdVals_nolake) %>% dplyr::filter(value >= 1) %>%
  summarise_at('count',sum)

# Total forest counts
forest_nolake = as.data.frame(nlcdVals_nolake) %>% dplyr::filter(value >= 41 & value <= 43) %>%
  summarise_at('count',sum)

perFor_nolake = forest_nolake/total_nolake
perFor_nolake

```

4.68% of Lake Mendota's catchment is forest if you exclude Lake Mendota from the total catchment area. If you include Lake Mendota in its catchment (which I would be inclined to do), the percent forest would be 4.39%

## Question 2: What is the area of forest in the Mendota catchment?

```{r, include=T}
res<-res(nlcd)
crs(nlcd) #check units
pixelarea<-res[1]*res[2] #in m2
forest_area <- forest*pixelarea/1000000
names(forest_area)[1]<-'area_km2'
print(forest_area)
```

The total area of forest in the Lake Mendota watershed is 26.4 km2



